---
title: "DataScienceProjectWeek2"
author: "Chia-Hsun Cheng"
date: "6/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(cache = TRUE)
```

## Get Corpora Data

```{r}
# https://www.coursera.org/learn/data-science-project/supplement/Iimbd/task-0-understanding-the-problem
zipfile = "./Coursera-SwiftKey.zip"
# download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip", zipfile)
# unzip(zipfile)

list.files("./final")
list.files("./final/en_US")
```

## Perform Exploratory Analysis on Corpora

## Analyze Line Count

```{bash}
wc -l ./final/en_US/en_US.twitter.txt
wc -l ./final/en_US/en_US.news.txt
wc -l ./final/en_US/en_US.blogs.txt
```

```{r}
library(fpeek)
options(scipen=999)
# If you cannot execute the RMarkdown generation for getting the error message 'no function peek_count_lines`, edit this block by adding spaces and execute again

getLineCounts <- function(filename) {
  return (peek_count_lines(filename)) # This is much faster
}
```

| File |Line Count | 
| :-: | :-: | 
| en_US.twitter.txt | `r getLineCounts("./final/en_US/en_US.twitter.txt")` |
| en_US.news.txt | `r getLineCounts("./final/en_US/en_US.news.txt")` |
| en_US.blogs.txt | `r getLineCounts("./final/en_US/en_US.blogs.txt")` |

### Define Tokenization

```{r, echo = FALSE, eval = FALSE}
library(stringr)
str_split(",1", "(?=[,])") # Buggy? So, we cannot use one run of str_split only
str_split("1,", "(?=[,])")
str_split("\"You", "(?=[\"])") # Buggy?
```

```{r}
library(stringr)

totokens <- function(line) {
  line <- gsub("’", "'", line)
  # https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_split
  tokens <- str_split(line, "([ ]|(?='m|'re)|(?=[ \\?\\.,!]))", simplify = TRUE)
  res <- c()
  for (token in tokens) {
    # For the above reason, we need to do strsplit again
    # https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/strsplit
    res <- c(res, unlist(strsplit(token, "(?=[,\"])", perl=TRUE)))
  }
  return (res)
}

totokens("I'm fine. How're you?")
totokens(",1")
totokens("1,")
totokens("\"This is a really excellent")
```

### Define Read/Sample Function

```{r}
readLineByPos <- function(filename, pos) {
    con <- file(filename, "r")
    line <- NULL
    for (i in 1:pos) {
      line <- readLines(con, n = 1)
      if (length(line) == 0) {
        break
      }
    }
    close(con)
    return (line)
}
line <- readLineByPos("./final/en_US/en_US.twitter.txt", 1)
line
totokens(line)
line <- readLineByPos("./final/en_US/en_US.news.txt", 757)
totokens(line)
```

Since the corpora is large, we only read a portion of it for efficiency.

We define `random_state` to read random lines of a file to avoid bias.

```{r}
library(dplyr)
# We define this function to cache read results
getReadFunc <- function(filename) {
  lineCounts <- getLineCounts(filename)
  con <- file(filename, "r")
  line <- readLines(con, n = -1)
  close(con)
  
  # Set `random` as TRUE for random
  # Set `random_state` as the random seed if random is toggled
  readFunc <- function(lineToRead = Inf, random_state = -1, random = TRUE) {
    if (is.infinite(lineToRead)) {
      lineToRead = lineCounts
    }
    
    inSample <- 1:lineToRead
    if (random) {
      if (random_state >= 0) {
          set.seed(random_state) 
      }
      inSample <- sample(lineCounts, lineToRead)
    }
    return (line[inSample])
  }
  return (readFunc)
}
readTwitter <- getReadFunc("./final/en_US/en_US.twitter.txt")
readTwitter(10,  random_state = 1)
readNews <- getReadFunc("./final/en_US/en_US.news.txt")
readBlogs <- getReadFunc("./final/en_US/en_US.blogs.txt")
```

### Analyize Number of Tokens in One Line

```{r}
getTokenFrame <- function(readFunc, readCount, random_state = -1) {
  lines <- readFunc(readCount,  random_state = random_state)
  df <- NULL
  for (line in lines) {
    tokens <- totokens(line)
    # We need to change the type of tokens to list here; otherwise, it would be flattened to several rows
    df <- bind_rows(df, list(a = line, b = list(tokens))) 
  }

  colnames(df) <- c("line", "tokens")
  return (df)
}

# readCount <- Inf
readCount <- 1000
dfTwitter <- getTokenFrame(readTwitter, readCount)
dfNews <- getTokenFrame(readNews, readCount)
dfBlogs <- getTokenFrame(readBlogs, readCount)
```

| |Read Line Count | Word Count |
| :-: | :-: | :-: |
| en_US.twitter.txt | `r length(dfTwitter$line)` | `r sum(sapply(dfTwitter$tokens, length))` |
| en_US.news.txt | `r length(dfNews$line)` | `r sum(sapply(dfNews$tokens, length))` |
| en_US.blogs.txt | `r length(dfBlogs$line)` | `r sum(sapply(dfBlogs$tokens, length))` |


```{r}
par(mfrow=c(3,1))
hist(sapply(dfTwitter$tokens, length), xlab="# tokens in one line", main="twitter")
abline(v=mean(sapply(dfTwitter$tokens, length)), col="blue")
hist(sapply(dfNews$tokens, length), xlab="# tokens in one line", main="news")
abline(v=mean(sapply(dfNews$tokens, length)), col="blue")
hist(sapply(dfBlogs$tokens, length), xlab="# tokens in one line", main="blogs")
abline(v=mean(sapply(dfBlogs$tokens, length)), col="blue")
```

If we analyze number of tokens in one line, we observe that blogs > news > twitter

## Get N-Gram

```{r}
# This function can be optimized
library(purrr)
library(tibble)
getNGramTable <- function(df, N, excludes = c(",", "!", "?", ".", "", " ", "-", "\"", "'"), coverage = 1) {
  res <- data.frame()
  cols <- c()
  for (i in 1:N) {
    colname <- paste('w', i, sep = "")
    res[1, colname] <- "?"
    cols <- c(cols, colname)
  }
  template <- as_tibble(res) # tibble is faster than data.frame
  l <- list()
  for (i in 1:nrow(df)) {
    curTokens <- unlist(df[i,]$tokens)
    now <- template
    # We skip for line with tokens less than N
    if (length(curTokens) < N) {
      next
    }
    for (k in 1:(length(curTokens)-N+1)) {
      segment <- curTokens[k:(k+N-1)]
      if (any(segment %in% excludes)) {
        next
      }
      now <- bind_rows(now, set_names(segment, colnames(res))) 
    }
    now <- slice(now, -1)
    l[[length(l)+1]] <- now
  }
  
  # It's faster to have several tibbles combine afterwards 
  res <- bind_rows(l)
  # We cannot use res[-1,] since for 1-dimension, it would become vector after transformation
  res <- res %>% slice(-1)
  nGramTable <- res %>%
    group_by_at(cols) %>%
    summarize(count = n()) %>%
    arrange(desc(count))
  
  total <- sum(nGramTable$count)
  nGramTable <- nGramTable %>%
    mutate(prob = count/total)
  
  # Remove based on coverage
  if (coverage < 1) {
    for (i in 1:nrow(nGramTable)) {
      if (sum(nGramTable$prob[1:i]) >= coverage) {
        nGramTable = nGramTable[1:i, ]
        break
      }
    } 
  }
  
  return (nGramTable)
}
head(getNGramTable(dfTwitter, 1))

# Performance test
# start_time <- Sys.time()
# getNGramTable(getTokenFrame(readTwitter, 1000, random_state = 1), 1)
# end_time <- Sys.time()
# end_time - start_time
```

We define the `getNGramTable` function to get N-gram from the data frame with probability.

### Word Freqeuncy

```{r}
table1Twitter <- getNGramTable(dfTwitter, 1)
table1News <- getNGramTable(dfNews, 1)
table1Blogs <- getNGramTable(dfBlogs, 1)
```

```{r}
par(mfrow=c(3,1))
plot(table(table1Twitter$count), main = "twitter", xlab = "# occurs per word", ylab = "count")
plot(table(table1News$count), main = "news", xlab = "# occurs per word", ylab = "count")
plot(table(table1Blogs$count), main = "blogs", xlab = "# occurs per word", ylab = "count")
```

```{r}
topN <- 5
ratio <- sum(table(table1Twitter$count)[1:topN])/sum(table(table1Twitter$count))
ratio
```

We observe that some words occur very frequently, while many of them occurs only seldom times.

Top `r topN` frequent words dominates `r ratio*100`% of all the counts. 


## Determine the Size of the Model

How do we determine how many lines to read for the corpora to make the model representable?

One criteria we adopted here is that if we read by two different `random_state` of the file, the model we built for the old model has an acceptable coverage on the new model.

```{r}
coverage <- 0.8
```

We use `r coverage` for coverage.

```{r}
linesToReadForModel <- 1000
for (readCount in seq(linesToReadForModel, 20000, by=100)) {
  existedTable <- getNGramTable(getTokenFrame(readTwitter, readCount, random_state = 1), 1)
  newTable <- getNGramTable(getTokenFrame(readTwitter, readCount, random_state = 2), 1)
  now <- sum((newTable$w1 %in% existedTable$w1)*newTable$count)/sum(newTable$count)
  # message("now: ", now)
  if (now > coverage) {
    linesToReadForModel <- readCount
    break
  }
}
linesToReadForModel
```

We use `r linesToReadForModel` lines for building our model for the target coverage `r coverage`.

## Most Frequent Words

After we define the coverage, we can go on having a basic understanding of more frequent N-grams in our corpora.

```{r}
dfTwitter <- getTokenFrame(readTwitter, linesToReadForModel)
dfNews <- getTokenFrame(readNews, linesToReadForModel)
dfBlogs <- getTokenFrame(readBlogs, linesToReadForModel)
table1Twitter <- getNGramTable(dfTwitter, 1)
table1News <- getNGramTable(dfNews, 1)
table1Blogs <- getNGramTable(dfBlogs, 1)
```

We build the 1-gram table for each text source and arrange by word frequency, so that we can observe frequent words in difference sources.

| | Frequent Words |
| :-: | :-: | :-: |
| en_US.twitter.txt | `r table1Twitter$w1[1:10]` |
| en_US.news.txt | `r table1News$w1[1:10]` | 
| en_US.blogs.txt | `r table1Blogs$w1[1:10]` | 

2-gram table

```{r}
table2Twitter <- getNGramTable(dfTwitter, 2)
table2News <- getNGramTable(dfNews, 2)
table2Blogs <- getNGramTable(dfBlogs, 2)
table2 <- rbind(table2Twitter, table2News, table2Blogs)
```

| | Frequent Words |
| :-: | :-: | :-: |
| en_US.twitter.txt | `r sapply(1:5, function(x) { paste(table2Twitter$w1[x], table2Twitter$w2[x]) })` |
| en_US.news.txt | `r sapply(1:5, function(x) { paste(table2News$w1[x], table2Twitter$w2[x]) })` | 
| en_US.blogs.txt | `r sapply(1:5, function(x) { paste(table2Blogs$w1[x], table2Twitter$w2[x]) })` | 

3-gram table

```{r}
table3Twitter <- getNGramTable(dfTwitter, 3)
table3News <- getNGramTable(dfNews, 3)
table3Blogs <- getNGramTable(dfBlogs, 3)
table3 <- rbind(table3Twitter, table3News, table3Blogs)
```

| | Frequent Words |
| :-: | :-: | :-: |
| en_US.twitter.txt | `r sapply(1:5, function(x) { paste(table3Twitter$w1[x], table3Twitter$w2[x], table3Twitter$w3[x])})` |
| en_US.news.txt | `r sapply(1:5, function(x) { paste(table3News$w1[x], table3News$w2[x], table3News$w3[x]) })` | 
| en_US.blogs.txt | `r sapply(1:5, function(x) { paste(table3Blogs$w1[x], table3Blogs$w2[x], table3Blogs$w3[x])})` | 

## Predict using N-gram Table

For a simple prediction algorithm, we can use matched N-1 grams for N grams table to predict the next word.

```{r}
getTableToUse <- function(tables, words) {
  for (table in tables) {
    numGram <- sum(grepl("w[0-9]+", colnames(table)))  
    if ((length(words)+1) == numGram) {
      return (table)
    }
  }
  return (NULL)
}

predict <- function(tables, words) {
  table <- getTableToUse(tables, words)
  if (is.null(table)) {
    stop("Unable to find table to use for length: ", length(words))
  }
  for (i in 1:length(words)) {
    # https://dplyr.tidyverse.org/reference/filter.html
    # https://stackoverflow.com/questions/46397148/dplyr-filter-by-the-first-column
    table <- table %>%
      filter_at(i,  all_vars(. == words[i]))
  }
  N = length(words)
  nxtCol = paste("w", N+1, sep="")
  table <- table %>%
       mutate(dprob = prob , nxt = .data[[nxtCol]]) 
  total <- sum(table$dprob)
  table$dprob <- table$dprob/total
    
  return (table)
}

# We need to pass list; otherwise, c(table1, table2) would combine tables
predict(list(table2, table3), c("in"))
predict(list(table2, table3), c("in", "the"))
```

* `nxt` is the predicted word
* `dprob` is the conditional probability under the N-1 gram

After we have the simple prediction function, we can apply it to develop our Shiny app.